Assumptions of Physics overview: \\ Classical mechanics and infinitesimal reducibility

### 2020-11-01 Overviews/Classical mechanics/ClassicalMechanicsOverview.tex
Minor edits & Intro paragraph

Now we generalize to multiple coordinates. We say two coordinates are independent if changing the units for one does not induce a change of units for the other. Now suppose our particle state space $\mathcal{S}$ is such that its units are fully defined by $n$ independent coordinates $q^i$. Suppose you change the first coordinate $q^1$ without changing the others. Then we will find a variable $k_1$ that changes as before so that the densities are invariant. Now change the second coordinate $q^2$ in the same way while also fixing $k_1$. Then we find a corresponding $k_2$. We can proceed in the same way until we exhaust all coordinates, which must also mean that there are no state variables left. We find that $\mathcal{S}$ is $2n$-dimensional and the state variables are $\xi^a = \{ q^i, k_i \}$. We define an independent degree of freedom as the space charted by a pair of such variables.

@@

Phase space (i.e.~symplectic manifold) is the only type of manifold that is able to support coordinate invariant distributions, which are required to describe an infinitesimally reducible system.

@@

The argument can also be constructed through statistical concepts (i.e.~determinism and reversibility means conservation of variance), thermodynamic concepts (i.e.~determinism and reversibility means only the state of the system is important for the evolution, the system is therefore isolated and must conserve energy) or information theoretic consideration (i.e.~under deterministic and reversible evolution the amount of information does not change, information entropy has to be conserved).

@@

First, as before, we apply the assumption to the particles, which means that for every evolution in phase space there should be one and only one trajectory. Note that each space variable $x^i$ is a coordinate, a state variable that defines a unit. In fact, the trajectories can be fully described by those units and only those units. So we can say that $q^i=x^i$, each $q^i$ will be paired with a conjugate $p_i$ and each state $\{q^i, p_i\}$ will be mapped to one and only one trajectory. At each point $x^i$, then, must pass infinitely many trajectories, one for each combination of $\{p_i\}$. Since the trajectories are differentiable in $x^i$, we can define a velocity $v^i = d_t x^i$. If the equations of motion were such that $v^i=v^i(q^i)$, then kinematic equivalence would fail as the full trajectory would be determined only by $q^i$. So we must have $v^i=v^i(q^i, p_i)$. The relationship must be invertible or kinematic equivalence would fail. At any given time, then, we must have the following relationship:
\begin{equation}
\begin{aligned}
x^i &= q^i \\
v^j &= d_t x^j = v^j(q^i, p_k)
\end{aligned}
\end{equation}

>>>>

Now we generalize to multiple coordinates. We say two coordinates are independent if changing the units for one does not induce a change of units for the other. Now suppose our particle state space $\mathcal{S}$ is such that its units are fully defined by $n$ independent coordinates $q^i$. Suppose you change the first coordinate $q^1$ without changing the others. Then we will find a variable $k_1$ that changes as before so that the densities are invariant. Now change the second coordinate $q^2$ in the same way while also fixing $k_1$. Then we find a corresponding $k_2$. We can proceed in the same way until we exhaust all coordinates, which must also mean that there are no state variables left. We find that $\mathcal{S}$ is $2n$-dimensional, and the state variables are $\xi^a = \{ q^i, k_i \}$. We define an independent degree of freedom as the space charted by a pair of such variables.

@@

Phase space (i.e.~a symplectic manifold) is the only type of manifold that is able to support coordinate invariant distributions, which are required to describe an infinitesimally reducible system.

@@

The argument can also be constructed through statistical concepts (i.e.~determinism and reversibility means conservation of variance), thermodynamic concepts (i.e.~determinism and reversibility means only the state of the system is important for the evolution; the system is therefore isolated and must conserve energy) or information theoretic consideration (i.e.~under deterministic and reversible evolution the amount of information does not change, so information entropy has to be conserved).

@@

First, as before, we apply the assumption to the particles, which means that for every evolution in phase space there should be one and only one trajectory. Note that each space variable $x^i$ is a coordinate, i.e. a state variable that defines a unit. In fact, the trajectories can be fully described by those units and only those units. So we can say that $q^i=x^i$, each $q^i$ will be paired with a conjugate $p_i$ and each state $\{q^i, p_i\}$ will be mapped to one and only one trajectory. At each point $x^i$, then, infinitely many trajectories must pass, one for each combination of $\{p_i\}$. Since the trajectories are differentiable in $x^i$, we can define a velocity $v^i = d_t x^i$. If the equations of motion were such that $v^i=v^i(q^i)$, then kinematic equivalence would fail as the full trajectory would be determined only by $q^i$. So we must have $v^i=v^i(q^i, p_i)$. The relationship must be invertible or kinematic equivalence would fail. At any given time, then, we must have the following relationship:
\begin{equation}
\begin{aligned}
x^i &= q^i \\
v^j &= d_t x^j = v^j(q^i, p_k)
\end{aligned}
\end{equation}
